---
title: "Invertebrate Abundance Wrangling & Statistics - Simple"
output: html_document
date: "2024-03-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Below is a simplified script of my invertebrate abundance data wrangling & analysis.

# Background 

Invertebrate abundance data was collected by tree-beating in July & August of 2023. 

# Loading packages and data

```{r}
library(AICcmodavg)
library(arm)
library(cowplot)
library(DAAG)
library(faraway)
library(ggfortify)
library(ggplot2)
library(lme4)
library(MASS)
library(nlme)
library(quarto)
library(SemiPar)
library(Sleuth3)
library(SMPracticals)
library(tidyverse)
library(dplyr)
library(lmerTest)
```

Loading in raw data:

```{r}
invert_data_raw <- read_csv("/Users/dorasolloway/Documents/Wadham/Year_4/Masters_project/invert_stats/TCC_beating.csv")
```

# Data wrangling

Cleaning names and filtering by species:

```{r}
invert_data_clean <- invert_data_raw %>% 
  clean_names() %>% 
  filter(tree_sp == "Oak"| tree_sp == "Alder" | tree_sp == "Birch") # The invertebrate dataset only includes oak, alder, birch and sitka spruce trees. My project focuses on broadleaves only so I will fitler out sitka
```

Adding in variables for basalt addition, microbiome innoculation and treatment - will do this by loading in another dataset and joining these columns by plot number. 

```{r}
# Reading in herbivory dataset with plot and treatment information 
herbivory_data <- read_csv("/Users/dorasolloway/Documents/Wadham/Year_4/Masters_project/Herbivory_stats/herbivory_data_raw.csv")

treatment_data <- herbivory_data %>% 
  select(plot, treatment, basalt_present, microbiome_present)

# Joining treatment data with invertebrate data
invert_treatment_data <- invert_data_clean %>% 
  select(block, plot, tree_sp, tree_ref, class, order, family, genus, species, abundance, juv, comment) %>% 
  left_join(treatment_data, by = "plot") 

# Removing repeated rows
invert_data_unique <- distinct(invert_treatment_data)
```

Filtering out observations with empty values for treatment (mistake in the data for the plot number):

```{r}
invert_data_unique <- invert_data_unique %>% 
  filter(!is.na(treatment))
```

# Converting variables to factors

```{r}
invert_data_unique$block <- factor(invert_data_unique$block)
invert_data_unique$plot <- factor(invert_data_unique$plot)
invert_data_unique$tree_sp <- factor(invert_data_unique$tree_sp)
invert_data_unique$tree_ref <- factor(invert_data_unique$tree_ref)
invert_data_unique$class <- factor(invert_data_unique$class)
invert_data_unique$order <- factor(invert_data_unique$order)
invert_data_unique$family <- factor(invert_data_unique$family)
invert_data_unique$genus <- factor(invert_data_unique$genus)
invert_data_unique$species <- factor(invert_data_unique$species)
invert_data_unique$juv <- factor(invert_data_unique$juv)
invert_data_unique$treatment <- factor(invert_data_unique$treatment)
invert_data_unique$basalt_present <- factor(invert_data_unique$basalt_present)
invert_data_unique$microbiome_present <- factor(invert_data_unique$microbiome_present)
invert_data_unique$abundance <- as.numeric(invert_data_unique$abundance)
```

# Model making, assumption checking and transformations

Will make a linear mixed effects model to determine the impact of the treatments on invertebrate abundance:

```{r}
invert_abundance_mem <- lmer(abundance ~ basalt_present * microbiome_present * tree_sp + 
                               (1|block) + (1|plot) + (1|tree_ref), 
                             data = invert_data_unique)
```

Checking the assumptions of the model: 

```{r}
plot(invert_abundance_mem)
qqnorm(resid(invert_abundance_mem))
qqline(resid(invert_abundance_mem))
```

The assumptions of normality is well met but not homogeneity of variance - extreme fan shape indicates data exhibits heteroscedasticity. Need to transform data.

```{r}
log_invert_mem <- lmer(log(abundance) ~ basalt_present * microbiome_present * tree_sp + 
                               (1|block) + (1|plot) + (1|tree_ref), data = invert_data_unique)
```

```{r}
plot(log_invert_mem)
qqnorm(resid(log_invert_mem))
qqline(resid(log_invert_mem))
```

This model is a much better fit for the residuals but now very non-normal. However, I will prioritise meeting the assumption of homogeneity of variance as linear mixed-effects models are normally more robust to violations of the assumption of normality. I will therefore log my residuals.

# Data analysis

```{r}
anova(log_invert_mem, type = "I")
```

There are no significnat effects of any of the treatments on invertebrate abundance. 

```{r}
summary(log_invert_mem)
```

# Obtaining summary statistics

I will use the marginal effects package to predict means and standard errors.

Loading the marginal effects package:

```{r}
library(marginaleffects)
```

Predicting summary statistics from the model grouped by treatment and species:

```{r}
log_predictions_invert_data <- predictions(log_invert_mem, by = c("microbiome_present", "basalt_present", "tree_sp"))
log_predictions_invert_data
```

I will now save the predictions dataframe:

```{r}
write_csv(log_predictions_invert_data, "log_predictions_invert_data.csv")
```

I will read the data back in to view the dataframe as formatting may have changed by saving as a csv file.

```{r}
log_predictions_invert_data <- read_csv("log_predictions_invert_data.csv")
log_predictions_invert_data
```

These predictions are logged so need to I need to back transform (exponentiate and -0.1) the estimates and confidence intervals to get actual estimates: 

```{r}
predictions_invert_data <- log_predictions_invert_data %>% 
  mutate(estimate = exp(estimate),
         conf.low = exp(conf.low),
         conf.high = exp(conf.high))
                  
predictions_invert_data
```

Saving the dataframe:

```{r}
#write_csv(predictions_invert_data, "predictions_invert_data.csv")
```

# Data visualisation

I will now load in the predictions dataframe to create graphs using ggplot.

```{r}
predictions_invert_data <- read_csv("predictions_invert_data.csv")
predictions_invert_data
```

Adding in a variable for treatment to make plotting easier:

```{r}
predictions_invert_data <- predictions_invert_data %>% 
  mutate(treatment = case_when(
    basalt_present == "y" & microbiome_present == "y" ~ "basalt+microbiome",
    basalt_present == "y" & microbiome_present == "n" ~ "basalt",
    basalt_present == "n" & microbiome_present == "y" ~ "microbiome",
    basalt_present == "n" & microbiome_present == "n" ~ "control",
    TRUE ~ NA_character_
  ))

predictions_invert_data
```

Need to reorder the factor levels within the dataset:

```{r}
predictions_invert_data$treatment <- factor(predictions_invert_data$treatment,
                                                levels = c("control", "basalt", "microbiome", "basalt+microbiome"))
```

Plotting predicted means and confidence intervals for decomposition rate under different treatments and for different species:

```{r}
predictions_invert_fig1 <- ggplot(predictions_invert_data, 
                                   aes(x = treatment, 
                                       y = estimate, 
                                       colour = treatment, group = treatment)) + 
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high), 
                  colour="black", width = 0.2, position = position_dodge(0.1)) +
    geom_point(position = position_dodge(0.1), size = 3) +
  scale_color_manual(values=c( "forestgreen", "darkorange", "dodgerblue", "deeppink"), 
                     labels=c("Control", "Basalt", "Microbiome", "Basalt+Microbiome")) +
  scale_x_discrete(labels = c( "C", "B", "M", "B+M")) +
  labs(x = "Treatment", y = "Invertebrate Abundance", title = "Effect of Basalt Addition and Microbiome\nInoculation on Invertebrate Abundance") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10),  
        axis.title.x = element_text(size = 14), 
        strip.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10),  
        axis.title.y = element_text(size = 14),  
        plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
        legend.title=element_blank()) +
  facet_wrap(~tree_sp, scales = "free_y", 
             labeller = labeller(species = c("alder" = "Alder", 
                                             "birch" = "Birch", 
                                             "oak" = "Oak")))

predictions_invert_fig1
```

*Figure 1: Graph showing the effect of basalt addition and microbiome inoculation on the predicted invertebrate abundnace for different species of tree and their confidence intervals, based on mixed-effects model 1*

The confidence intervals on this graph are very large, indicating that there are no significant effects of the treatments on invertebrate abundance.

# Obtaining estimates for treatments overall (not by species)

```{r}
log_predictions_invert_data_overall <- predictions(log_invert_mem, by = c("microbiome_present", "basalt_present"))
log_predictions_invert_data_overall
```

I will now save this dataframe:

```{r}
#write_csv(log_predictions_invert_data_overall, "log_predictions_invert_data_overall.csv")
```

Reading in the dataframe:

```{r}
log_predictions_invert_data_overall <- read_csv("log_predictions_invert_data_overall.csv")
log_predictions_invert_data_overall
```

These predictions are logged so need to exponentiate and -0.1 the estimates and confidence intervalues to get actual values:

```{r}
predictions_invert_data_overall <- log_predictions_invert_data_overall %>% 
  mutate(estimate = exp(estimate), 
         conf.low = exp(conf.low),
         conf.high = exp(conf.high))
                  
predictions_invert_data_overall
```

Saving the dataframe:

```{r}
#write_csv(predictions_invert_data_overall, "predictions_invert_data_overall.csv")
```

I will now load in the predictions dataframe to create graphs using ggplot.

```{r}
predictions_invert_data_overall <- read_csv("predictions_invert_data_overall.csv")
```

Adding in a variable for treatment to make plotting easier:

```{r}
predictions_invert_data_overall <- predictions_invert_data_overall %>% 
  mutate(treatment = case_when(
    basalt_present == "y" & microbiome_present == "y" ~ "basalt+microbiome",
    basalt_present == "y" & microbiome_present == "n" ~ "basalt",
    basalt_present == "n" & microbiome_present == "y" ~ "microbiome",
    basalt_present == "n" & microbiome_present == "n" ~ "control",
    TRUE ~ NA_character_
  ))

predictions_invert_data_overall
```

Need to reorder the factor levels within the dataset:

```{r}
predictions_invert_data_overall$treatment <- factor(predictions_invert_data_overall$treatment,
                                                levels = c("control", "basalt", "microbiome", "basalt+microbiome"))
```

Plotting predicted means and confidence intervals for decomposition rate under different treatments and for different species:

```{r}
predictions_invert_fig2 <- ggplot(predictions_invert_data_overall, 
                                   aes(x = treatment, 
                                       y = estimate, 
                                       colour = treatment, group = treatment)) + 
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high), 
                  colour="black", width=.3, linewidth= 0.5, position = position_dodge(0.1)) +
    geom_point(position = position_dodge(0.1), size=4) +
  scale_color_manual(values=c( "forestgreen", "darkorange", "dodgerblue", "deeppink"), 
                     labels=c("Control", "Basalt", "Microbiome", "Basalt+Microbiome")) +
  scale_x_discrete(labels = c( "C", "B", "M", "B+M")) +
  labs(x = "Treatment", y = "Invertebrate Abundance", title = "Invertebrate Abundance") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10),  
        axis.title.x = element_text(size = 14), 
        strip.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10),  
        axis.title.y = element_text(size = 14),  
        plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
        aspect.ratio = 0.9,
        legend.title=element_blank()) 

predictions_invert_fig2
```

*Figure 2: Graph showing the effect of basalt addition and microbiome inoculation on the overall predicted invertebrate abundance, and their confidence intervals, based on mixed-effects model 1*

From the above graph we can see that overall, invertebrate abundance is not significantly affected by treatment.

# Saving the plots

```{r}
#ggsave("predictions_invert_fig1.png", plot = predictions_invert_fig1, width = 9, height = 4)
```

```{r}
#ggsave("predictions_invert_fig2.png", plot = predictions_invert_fig2, width = 8, height = 6)
```

# Obtaining microbiome only summary statistics

```{r}
log_predictions_invert_data_microbiome_overall <- predictions(log_invert_mem, by = c("microbiome_present"))
log_predictions_invert_data_microbiome_overall
```

Back-transforming estimates and confidence intervals:

```{r}
predictions_invert_data_microbiome_overall <- log_predictions_invert_data_microbiome_overall %>% 
  mutate(estimate = exp(estimate), 
         conf.low = exp(conf.low),
         conf.high = exp(conf.high))
                  
predictions_invert_data_microbiome_overall
```

Species-by-species predictions

```{r}
log_predictions_invert_data_microbiome <- predictions(log_invert_mem, by = c("microbiome_present", "tree_sp"))
```

Back-transforming estimates and confidence intervals:

```{r}
predictions_invert_data_microbiome <- log_predictions_invert_data_microbiome %>% 
  mutate(estimate = exp(estimate), 
         conf.low = exp(conf.low),
         conf.high = exp(conf.high))
                  
predictions_invert_data_microbiome
```

# Obtaining basalt only summary statistics

```{r}
log_predictions_invert_data_basalt_overall <- predictions(log_invert_mem, by = c("basalt_present"))
log_predictions_invert_data_basalt_overall
```

```{r}
predictions_invert_data_basalt_overall <- log_predictions_invert_data_basalt_overall %>% 
  mutate(estimate = exp(estimate), 
         conf.low = exp(conf.low),
         conf.high = exp(conf.high))
                  
predictions_invert_data_basalt_overall
```

Species-by-species predictions

```{r}
log_predictions_invert_data_basalt <- predictions(log_invert_mem, by = c("basalt_present", "tree_sp"))
log_predictions_invert_data_basalt %>% 
  arrange(tree_sp)
```

Back-transforming estimates and confidence intervals:

```{r}
predictions_invert_data_basalt <- log_predictions_invert_data_basalt %>% 
  mutate(estimate = exp(estimate), 
         conf.low = exp(conf.low),
         conf.high = exp(conf.high))
                  
predictions_invert_data_basalt
```







